---
title: "CoreGx: Class and Function Abstractions for PharmacoGx, RadioGx and ToxicoGx"

author: 
- 'Petr Smirnov'
- 'Ian Smith'
- 'Christopher Eeles'
- 'Benjamin Haibe-Kains'
  
output: pdf_document
  
vignette: |
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{CoreGx: Class and Function Abstractions}
---
# CoreGx

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

This package provides a foundation for the PharmacoGx, RadioGx and ToxicoGx 
packages. It is not intended for standalone use, only as a dependency for the 
aforementioned software. Its existence allows abstracting generic definitions, 
method definitions and class structures common to all three of the Gx suite 
packages. Examples here are included as a demonstration only.

```{r coregx_load}
library(CoreGx)
```


```{r coregx_do_some_stuff}
data(clevelandSmall_cSet)

# Setter method for sample phenotype information
phenoInfo(clevelandSmall_cSet, 'rna') <- phenoInfo(clevelandSmall_cSet, 'rna')

# Setter method for molecular features available in the assay samples
featureInfo(clevelandSmall_cSet, 'rna') <- featureInfo(clevelandSmall_cSet, 'rna')

```

For more information about this package and the possibility of collaborating on 
its extension please contact benjamin.haibe.kains@utoronto.ca.