---
title: "CoreGx Vignette"
author: "Benjamin Haibe-Kains"
date: "25/03/2020"
output: html_document
---
# CoreGx

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This package provides a foundation for the PharmacoGx, RadioGx and ToxicoGx packages. It is not intended for standalone use, only as a dependency for the aforementioned software. Its existence allows abstracting generic definitions, method definitions and class structures common to all three of the Gx suite packages. Examples here are included as a demonstration only.

```{r coregx_load}
library(CoreGx)
```


```{r coregx_do_some_stuff}
data(clevelandSmall)

# Setter method for sample phenotype information
phenoInfo(clevelandSmall, 'rna') <- phenoInfo(clevelandSmall, 'rna')

# Setter method for molecular features available in the assay samples
featureInfo(clevelandSmall, 'rna') <- featureInfo(clevelandSmall, 'rna')

```

For more information about this package and the possibility of collaborating on its extension please contacnt benjamin.haibe.kains@utoronto.ca.